Diffusive Homeostasis in a Self-Organizing Recurrent Neural Network \\ \begin{large} Spatially dependent Interaction as a Determinant of Neural Activity and Plasticity \end{large}

### 2016-12-10 draft.tex
.

After the elimination of oscillatory activity in the network, we further investigated the features of spiking behaviour, in particular those of the excitatory population, since these were exposed to homeostasis.

Balanced networks of inhibitory and excitatory populations of spiking LIF-neurons are known to show highly random, uncorrelated spiking behaviour under sparse connectivity with a relatively high per-synapse connection strength, see e.g. \cite{Vreeswijk1996} or \cite{Brunel2000}. This kind of asynchronous regular state is characterized by a population firing rate showing no regularities over time, though maintaining an overall stable level of activity. This feature was shown in the previous section after slowing down the rate of homeostatic adaption. On the level of single cells, interspike intervals should be random, approximately following the statistics of a poisson process.   
\begin{figure}
\includegraphics[width=\textwidth]{../../plots/Spike_Stats/isi_cv_compare.png}
\caption{A/C: Distribution of interspike intervals of excitatory/inhibitory neurons for diffusive and non-diffusive homeostasis, measured over $\mathrm{1000\,s}$. Only neurons with a mean firing rate in a certain range (see legend) were included into statistics to prevent overlap of different firing rates (except for non-diff. homeostasis and excitatory neurons, all being very close to 3 Hz). Black lines are fits of exponential functions. B/D: Distribution of coefficients of variation. Each sample represents the CV of one excitatory/inhibitory neuron.}
\label{ISI_CV_compare}
\end{figure}
Figure \ref{ISI_CV_compare} shows the distribution of interspike intervals (ISIs) and the distribution of coefficient of variation ($\mathrm{CV\equiv\sigma/\mu}$), taken over the set of excitatory and inhibitory neurons. However, in order to provide comparability between the ISI distributions in the diffusive/non-diffusive case, we restricted the statistics of ISIs to neurons whose mean firing rate fall into a window with a width of 1 Hz and a certain mean (see legend in Figure \ref{ISI_CV_compare}). Apparently, an approximately exponential distribution of ISIs is preserved for diffusive homeostasis (represented by straight lines in a log-plot), which is characteristic for poisson processes \cite[p. 27]{Theor_Neur_Dayan}. For perfect poissonian spiking, the CV should give a value of 1. CVs that are smaller/larger, as seen in Figure \ref{ISI_CV_compare} B/D, indicate a tendency towards greater/less regularity. This is presumably due to the statistics of ISIs close to zero: Interestingly, while excitatory neurons have an  under-representation of short ISIs, the opposite is the case for inhibitory neurons. These types of behaviour of LIF-models have been shown to depend on the CV of input the neuron receives \cite{Ostojic2011}: A smaller signal-to-noise-ratio (larger CV) of neural input leads to a larger CV of ISIs for the output and an over-representation of small intervals, see Figure \ref{Ostojic_ISI_plot}.
\begin{figure}
\includegraphics[width=\textwidth]{../../plots/external_plots/ostojic_isi.png}
\caption{Plot was taken from \cite{Ostojic2011}. Top row: sample trajectories of the membrane potential of a LIF-neuron (blue curve) and the corresponding probability density (yellow). From left to right, the input's CV is successively increased. Bottom row: the corresponding ISI distributions and the dominating exponential curve for large ISIs (red).}
\label{Ostojic_ISI_plot}
\end{figure}
This relation fits nicely into our set of network parameters: inhibitory neurons receive stronger recurrent inhibitory input ($\mathrm{-60\,mV}$ in total) compared to excitatory neurons ($\mathrm{-12\,mV}$). This lowers the signal-to-noise ratio for the inputs of inhibitory neurons, leading to the aforementioned statistical differences. Still, one can observe a dominant exponential component for both excitatory and inhibitory neurons.

>>>>

After the elimination of oscillatory activity in the network, we further investigated the features of neural activity, in particular of the excitatory population, since it was exposed to homeostasis.

\subsubsection{Spike-Train Statistics}
Balanced networks of inhibitory and excitatory populations of spiking LIF-neurons are known to show highly random, uncorrelated spiking behaviour under sparse connectivity with a relatively high per-synapse connection strength, see e.g. \cite{Vreeswijk1996} or \cite{Brunel2000}. This kind of asynchronous regular state is characterized by a population firing rate showing no regularities over time, though maintaining an overall stable level of activity. This feature was shown in the previous section after slowing down the rate of homeostatic adaption. On the level of single cells, interspike intervals should be random, approximately following the statistics of a poisson process.   
\begin{figure}
\includegraphics[width=\textwidth]{../../plots/Spike_Stats/isi_cv_compare.png}
\caption{A/C: Distribution of interspike intervals of excitatory/inhibitory neurons for diffusive and non-diffusive homeostasis, measured over $\mathrm{1000\,s}$. Only neurons with a mean firing rate in a certain range (see legend) were included into statistics to prevent overlap of different firing rates (except for non-diff. homeostasis and excitatory neurons, all being very close to 3 Hz). Black lines are fits of exponential functions. B/D: Distribution of coefficients of variation. Each sample represents the CV of one excitatory/inhibitory neuron.}
\label{ISI_CV_compare}
\end{figure}
Figure \ref{ISI_CV_compare} shows the distribution of interspike intervals (ISIs) and the distribution of coefficient of variation ($\mathrm{CV\equiv\sigma/\mu}$), taken over the set of excitatory and inhibitory neurons. However, in order to provide comparability between the ISI distributions in the diffusive/non-diffusive case, we restricted the statistics of ISIs to neurons whose mean firing rate fall into a window with a width of 1 Hz and a certain mean (see legend in Figure \ref{ISI_CV_compare}). Apparently, an approximately exponential distribution of ISIs is preserved for diffusive homeostasis (represented by straight lines in a log-plot), which is characteristic for poisson processes \cite[p. 27]{Theor_Neur_Dayan}. For perfect poissonian spiking, the CV should give a value of 1. CVs that are smaller/larger, as seen in Figure \ref{ISI_CV_compare} B/D, indicate a tendency towards greater/less regularity. This is presumably due to the statistics of ISIs close to zero: Interestingly, while excitatory neurons have an  under-representation of short ISIs, the opposite is the case for inhibitory neurons. These types of behaviour of LIF-models have been shown to depend on the CV of input the neuron receives \cite{Ostojic2011}: A smaller signal-to-noise-ratio (larger CV) of neural input leads to a larger CV of ISIs for the output and an over-representation of small intervals. This relation fits nicely into our set of network parameters: inhibitory neurons receive stronger recurrent inhibitory input ($\mathrm{-60\,mV}$ in total) compared to excitatory neurons ($\mathrm{-12\,mV}$). This lowers the signal-to-noise ratio for the inputs of inhibitory neurons, leading to the aforementioned statistical differences. Still, one can observe a dominant exponential component for both excitatory and inhibitory neurons.

As an additional possibility of spike-train analysis, we looked at potential correlations across neurons within both populations. To do so, we calculated cross-correlations for all possible neuron-pairings within a population, excluding auto-correlations. Figure \ref{Pop_Mean_Cross_Corr} shows the mean cross-correlation taken over all these pairs. Surprisingly, we found a significant increase in correlation across neurons for diffusive homeostasis. 
\begin{figure}
\includegraphics[width=\textwidth]{../../plots/Spike_Stats/pop_mean_cross_corr.png}
\caption{Mean of cross correlation over all pairs of neurons in the respective neuron-group. Significantly higher correlation can be found in both groups for diffusive homeostasis. Spike data was taken from $\mathrm{1400-1500\,s}$ of the simulation.}
\label{Pop_Mean_Cross_Corr}
\end{figure}
When speculating about the root of this discrepancy one could consider two possibilities: either, it is an immediate consequence of diffusive homeostasis, or it is indirectly related through other mechanisms present within the network. Regarding the first, it is certainly true that some amount of information about the activity of one neuron is transferred to another one in its neighbourhood via diffusion. However, due to the time scale of the homeostatic adaption, the response to fluctuations in the diffusive signal is on the order of seconds, while correlations in Figure \ref{Pop_Mean_Cross_Corr} are taking place within milliseconds. Thus, we concluded that a \emph{direct} link between homeostatic signalling and increased correlations can be regarded as implausible and that it must be an effect based on topological differences within the network. We investigate these in Section (???).

In summary, we found that diffusive homeostasis does not interfere with the poisson-like spiking statistics observed on a single-cell level, but leads to an increased correlation across the neural population.

\subsubsection{Distribution of Firing Rates}

!!!HEAD!!!
